Preliminary insights into the use of pottery and culinary practices at Guijiabao site in southwest China

### 2024-05-24 analysis/paper/paper.Rmd
udpated code for figures

# plot the first group
chroma_four <-
  meth_all %>%
  filter(name %in% c("SYG-3","SYG-TN13-E22-2#1","SYG-TN13-E22-1","SYG-TN13-E23-3","CDG-062")) %>%
  mutate(File = ifelse(File > 1000000, 1000000, File)) %>%
  ggplot(aes(Path, File)) +
  geom_line(size = 0.3) +
  ggrepel::geom_text_repel(data = highlight_common, #ggrepel::geom_text_repel
            aes(label = FA, angle = 0),
            size = 3,
            nudge_y = 115000,
            show.legend = FALSE,
            parse = TRUE) +
  geom_text(data = highlight_IS,
            aes(label = FA),
            size = 3,
            nudge_y = 100000,
            show.legend = FALSE) +
  geom_segment(aes(x= 41.1, xend= 41.5, y= 1000000, yend= 1000000), size =0.3) +
  geom_segment(aes(x= 41.1, xend= 41.5, y= 1030000, yend= 1030000), size =0.3) +
  annotate("text", x = 41.3, y = 1100000, label = "IS", size =3) +
  facet_wrap(~name, ncol = 1) +
  scale_y_continuous(limits = c(0, 1160000),
                     breaks = seq(0, 1100000, 200000),
                     expand = c(0, 0)) +
  scale_x_continuous(limits = c(18, 47),
                     breaks = seq(15, 50, 1),
                     labels = insert_minor(seq(15, 50, 5), 4),
                     expand = c(0, 0)) +
  labs(x = "Retention time (mins)", y = expression("Relative abundance " %->% "")) +
  theme_classic() +
  theme(axis.line.y = element_blank(), strip.background = element_blank(), # remove title box
        axis.ticks.y=element_blank(), axis.text.y = element_blank())

ggsave(plot = chroma_four,
       here::here("analysis","figures", "chromatograms_four.png"),
       bg = "white",
       width = 6,
       height = 8,
       dpi = 300,
       units = "in")
```

@@

delta_meth_data <-
  delta_meth_cor %>%
  dplyr::mutate(delta = `13C C18:0`-`13C C16:0`) %>%
  mutate(name = case_when(C13 %in% c("Pot 1") ~ "CDG-062",
                          C13 %in% c("Pot 2") ~ "SYG-3",
                          C13 %in% c("Pot 3") ~ "SYG-TN13-E22-2#1",
                          C13 %in% c("Pot 4") ~ "SYG-TN13-E22-1",
                          C13 %in% c("Pot 5") ~ "SYG-TN13-E22-2#2",
                          C13 %in% c("Pot 6") ~ "SYG-TN13-E23-3")) %>%
  mutate(name = factor(name, levels = c("SYG-TN13-E22-1",
                                        "SYG-TN13-E22-2#1",
                                        "SYG-TN13-E22-2#2",
                                        "SYG-TN13-E23-3",
                                        "SYG-3",
                                        "CDG-062"))) %>%
  mutate(period = ifelse(name %in% c("SYG-TN13-E22-1", "SYG-3" ), "Historical", "Neolithic"))

@@

# plot the carbon isotopes 
isotope_C16_C18 <-
  delta_meth_data %>% #delta_meth_cor/delta_blank_cor
  ggplot(aes(`13C C16:0`,`13C C18:0`)) +
  geom_point(size = 2, alpha = 0.9, aes(shape = name, color = period)) +
  #ggrepel::geom_text_repel(aes(label = C13)) +
  labs(x = expression(paste(delta*{}^13* C[16:0], "\u2030")),
       y = expression(paste(delta*{}^13* C[18:0], "\u2030"))) +
  scale_x_continuous(limits = c(-33, -20), breaks = seq(-33, -20, 2)) +
  scale_y_continuous(limits = c(-33, -20), breaks = seq(-33, -20, 2)) +#coord_fixed(ratio = 1)
  geom_abline(intercept = -0.7, linetype = "dashed") + #Suryanarayan et al.2021
  geom_abline(intercept = -3.1, linetype = "dashed") +
  annotate("text", x = -21.5, y = -22, angle = 45, vjust = 1.5,
           label = bquote(Delta*{}^13*"C = -0.7 \u2030")) +
  annotate("text", x = -21.5, y = -24.5, angle = 45, vjust = 1.5,
           label = bquote(Delta*{}^13*"C = -3.1 \u2030")) +
  annotate("text", x = c(-26, -27), y = c(-23.5, -32),
           label = c("Non-Ruminant fat", "Ruminant fat")) +
  annotation_raster(tem3, ymin = -38, ymax= -17.9, xmin = -36, xmax = -18.3) +
  theme_minimal()

big_delta_C16 <-
  delta_meth_data %>%
  ggplot(aes(`13C C16:0`, delta)) +
  geom_point(size = 2, alpha = 0.9, aes(shape = name, color = period)) +
  labs(x = expression(paste(delta*{}^13* C[16:0], "\u2030")),
       y = expression(paste(delta*{}^13* C[18:0], "\u2030"))) +
  scale_x_continuous(limits = c(-33, -20), breaks = seq(-33, -20, 2)) +
  ylim(-3, 1.5) +
  geom_hline(yintercept = -0.7, linetype = "dashed") +
  annotate("text", x = -31.5, y = -0.9,
           label = bquote(Delta*{}^13*"C = -0.7 \u2030")) +
  annotate("text", x = c(-31, -21), y = c(1.3, 1.3),
           label = c("C[3]~diet", "C[4]~diet"), parse= TRUE) +
  annotate("segment", x = -29, y = 1.3, xend = -23, yend = 1.3,
           arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
  annotate("text", x = c(-22, -21.5), y = c(-1.1, -0.3),
           label = c("Ruminant", "Non-ruminant")) +
  theme_minimal()

>>>>

# plot the first group
chroma_five <-
  meth_all %>%
  filter(name %in% c("SYG-3","SYG-TN13-E22-2#1","SYG-TN13-E22-1","SYG-TN13-E23-3","CDG-062")) %>%
  mutate(File = ifelse(File > 1000000, 1000000, File)) %>%
  ggplot(aes(Path, File)) +
  geom_line(size = 0.3) +
  ggrepel::geom_text_repel(data = highlight_common, #ggrepel::geom_text_repel
            aes(label = FA, angle = 0),
            size = 3,
            nudge_y = 115000,
            show.legend = FALSE,
            parse = TRUE) +
  geom_text(data = highlight_IS,
            aes(label = FA),
            size = 3,
            nudge_y = 100000,
            show.legend = FALSE) +
  geom_segment(aes(x= 41.1, xend= 41.5, y= 1000000, yend= 1000000), size =0.3) +
  geom_segment(aes(x= 41.1, xend= 41.5, y= 1030000, yend= 1030000), size =0.3) +
  annotate("text", x = 41.3, y = 1100000, label = "IS", size =3) +
  facet_wrap(~name, ncol = 1) +
  scale_y_continuous(limits = c(0, 1160000),
                     breaks = seq(0, 1100000, 200000),
                     expand = c(0, 0)) +
  scale_x_continuous(limits = c(17, 48),
                     breaks = seq(15, 50, 1),
                     labels = insert_minor(seq(15, 50, 5), 4),
                     expand = c(0, 0)) +
  labs(x = "Retention time (mins)", y = expression("Relative abundance " %->% "")) +
  theme_classic() +
  theme(axis.line.y = element_blank(), strip.background = element_blank(), # remove title box
        axis.ticks.y=element_blank(), axis.text.y = element_blank())

ggsave(plot = chroma_five,
       here::here("analysis","figures", "chromatograms_five.png"),
       bg = "white",
       width = 6,
       height = 8,
       dpi = 300,
       units = "in")
```

@@

delta_meth_data <-
  delta_meth_cor %>%
  dplyr::mutate(delta = `13C C18:0`-`13C C16:0`) %>%
  mutate(label = case_when(C13 %in% c("Pot 1") ~ "CDG-062",
                          C13 %in% c("Pot 2") ~ "SYG-3",
                          C13 %in% c("Pot 3") ~ "SYG-TN13-E22-2#1",
                          C13 %in% c("Pot 4") ~ "SYG-TN13-E22-1",
                          C13 %in% c("Pot 5") ~ "SYG-TN13-E22-2#2",
                          C13 %in% c("Pot 6") ~ "SYG-TN13-E23-3")) %>%
  mutate(label = factor(label, levels = c("SYG-TN13-E22-1",
                                        "SYG-TN13-E22-2#1",
                                        "SYG-TN13-E22-2#2",
                                        "SYG-TN13-E23-3",
                                        "SYG-3",
                                        "CDG-062"))) %>%
  mutate(period = ifelse(label %in% c("SYG-TN13-E22-1", "SYG-3" ), "Historical", "Neolithic"))

@@

# plot the carbon isotopes 
isotope_C16_C18 <-
  delta_meth_data %>% #delta_meth_cor/delta_blank_cor
  ggplot(aes(`13C C16:0`,`13C C18:0`)) +
  geom_point(size = 2, alpha = 0.9, aes(shape = label, color = period)) +
  #ggrepel::geom_text_repel(aes(label = C13)) +
  labs(x = expression(paste(delta*{}^13* C[16:0], "\u2030")),
       y = expression(paste(delta*{}^13* C[18:0], "\u2030"))) +
  scale_x_continuous(limits = c(-33, -20), breaks = seq(-33, -20, 2)) +
  scale_y_continuous(limits = c(-33, -20), breaks = seq(-33, -20, 2)) +#coord_fixed(ratio = 1)
  geom_abline(intercept = -0.7, linetype = "dashed") + #Suryanarayan et al.2021
  geom_abline(intercept = -3.1, linetype = "dashed") +
  annotate("text", x = -21.5, y = -22, angle = 45, vjust = 1.5,
           label = bquote(Delta*{}^13*"C = -0.7 \u2030")) +
  annotate("text", x = -21.5, y = -24.5, angle = 45, vjust = 1.5,
           label = bquote(Delta*{}^13*"C = -3.1 \u2030")) +
  annotate("text", x = c(-26, -27), y = c(-23.5, -32),
           label = c("Non-Ruminant fat", "Ruminant fat")) +
  annotation_raster(tem3, ymin = -38, ymax= -17.9, xmin = -36, xmax = -18.3) +
  theme_minimal()

big_delta_C16 <-
  delta_meth_data %>%
  ggplot(aes(`13C C16:0`, delta)) +
  geom_point(size = 2, alpha = 0.9, aes(shape = label, color = period)) +
  labs(x = bquote(delta*{}^13*"C 16:0 \u2030"),
       y = bquote(Delta*{}^13*"C")) +
  scale_x_continuous(limits = c(-33, -20), breaks = seq(-33, -20, 2)) +
  ylim(-3, 1.5) +
  geom_hline(yintercept = -0.7, linetype = "dashed") +
  annotate("text", x = -31.5, y = -0.9,
           label = bquote(Delta*{}^13*"C = -0.7 \u2030")) +
  annotate("text", x = c(-31, -21), y = c(1.3, 1.3),
           label = c("C[3]~diet", "C[4]~diet"), parse= TRUE) +
  annotate("segment", x = -29, y = 1.3, xend = -23, yend = 1.3,
           arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
  annotate("text", x = c(-22, -21.5), y = c(-1.1, -0.3),
           label = c("Ruminant", "Non-ruminant")) +
  theme_minimal()
